---
title: "Data Collection and Data Wrangling"
author: "Abdul Majid"
date: "February 1, 2022"
output:
  pdf_document: default
  html_document: default
---

##### NOTE: 
1. Change author name and date to your exercise submission date in above section
2. Your code MUST execute without errors. 
3. You can add more lines in your code as required.

## Section 1: Manipulating Data

### Question 1 
**The dataset is loaded for you. Perform the following tasks:**  
1. use the USPersonalExpenditure dataset and store in a variable.  
2. Compute summary statistics: mean, standard deviation (sd) across time for each of the indicators.  
3. Create a new column that contains average values for each of the indicators.  


```{r }
library(help='datasets')
```


```{r }


```

```{r }
### start solution ####
```


```{r }
USPersonalExpenditure
var_dataset <- as.data.frame(USPersonalExpenditure)
for (i in seq_along(1:ncol(USPersonalExpenditure)))
{
  avg <- mean(var_dataset[,i])
  print(paste('Mean of', colnames(USPersonalExpenditure)[i], ' = ', avg))
  s_deviation <- sd(var_dataset[,i])
  print(paste('Standard Deviation of ', colnames(USPersonalExpenditure)[i], ' = ', s_deviation))
}
```


```{r }

mean_rows <- rowMeans(var_dataset)

var_dataset$'Average' <- mean_rows
var_dataset
```


```{r }

```


```{r }
### end solution ####
```

### Question 2 
**download the data from the available URL:**  
1. Create a new column containing the average bodywt (body weight) of each genus and of each order.  
2. Create a dataframe containing the average REM sleep time for each order.  
3. How many missing values are there (in total) and per column?  
4. How would you like to impute missing values? Write Justification.  <i> Hint: Overall Mean/median/mode vs. Groupby Mean/median/mode?


```{r }
url <- "https://raw.githubusercontent.com/genomicsclass/dagdata/master/inst/extdata/msleep_ggplot2.csv"



if(!file.exists("msleep_ggplot2.csv"))
{
  download.file(url, destfile = "msleep_ggplot2.csv",method = 'libcurl')
}

r <- read.csv("msleep_ggplot2.csv")
df <- data.frame(r)

avg_bodywt_genus <- aggregate(df$bodywt, list(df$genus), mean)
avg_bodywt_genus

avg_bodywt_order <- aggregate(df$bodywt, list(df$order), mean)
avg_bodywt_order

avg_sleepREM_order <- aggregate(df$sleep_rem, list(df$order), mean)
avg_sleepREM_order <- as.data.frame(avg_sleepREM_order)
avg_sleepREM_order

print(paste("Total NA(s) in whole file = ", sum(is.na(df))))

for (j in seq_along(1:ncol(df)))
{

  print(paste("NA(s) in column ", j, " = ", sum(is.na(df[, j]))))
}
df

```


```{r }
### start solution ####
```


```{r }

### For a missing value of a certain group, we will use mean of that group.
### Otherwise if there is a group of only one specie then we will use the overall mean

```


```{r }

```


```{r }

```


```{r }
### end solution ###
```

### Good job! You've completed this section!

## Section 2: Tidyverse

### Question 1
**Use the above dataset and perform the following tasks using any library from tidyverse:**  
1. Filter results to print average REM sleep and average total sleep  for those animals who are carnivores and then for those who are primates.  
2. Use the order column and "spread" it across the rest of the observations.


```{r }
### start solution ####
```


```{r }

library('dplyr')
library('tidyr')
df

```


```{r }

df %>% 
  filter(order=="Primates") %>% 
  select(order, sleep_total, sleep_rem)

```


```{r }

df %>% 
  filter(order=="Carnivora") %>% 
  select(order, sleep_total, sleep_rem)

```


```{r }

wide_DF <- df %>% spread(order, sleep_total) %>% select(-name, -genus, -vore, -conservation,
                                                        -sleep_rem, -awake, -sleep_cycle, 
                                                        -brainwt, -bodywt)
head(wide_DF, 24)

```


```{r }

```


```{r }

```


```{r }
### end solution ####
```

### Good job! You've completed this entire exercise!
